---
Author: Jace Higa
Date: Janurary 29th, 2025
---
**Abstract:**

This is a technical blog post of **both** an HTML file *and* [.qmd file](src/wine_features.qmd) hosted on GitHub pages.


**Step Up Code:**
```{r}
library(tidyverse)
library(caret)
library(fastDummies)
wine <- readRDS(gzcon(url("https://github.com/cd-public/D505/raw/master/dat/wine.rds")))
```

**Explanataion:**

> <span style="color:red;font-weight:bold">TODO</span>: *This code renders TODO into a red and bolded version of it.*

# Feature Engineering

We begin by engineering an number of features.

1. Create a total of 10 features (including points). 
2. Remove all rows with a missing value. 
3. Ensure only log(price) and engineering features are the only columns that remain in the `wino` dataframe.

```{r}
wino <- wine %>% 
  mutate(points = fct_lump(factor(points), 5)) %>%
  mutate(country = fct_lump(country, 4)) %>%
  mutate(lprice=log(price)) %>%
  select(lprice, points, country) %>%
  drop_na(.)
```

# Caret

We now use a train/test split to evaluate the features.

1. Use the Caret library to partition the wino dataframe into an 80/20 split. 
2. Run a linear regression with bootstrap resampling. 
3. Report RMSE on the test partition of the data.

```{r}
# 80/20 Split
wine_index <- createDataPartition(wino$lprice, p = 0.8, list = FALSE)
wino_tr <- wino[wine_index, ]
wino_te <- wino[-wine_index, ]
```


```{r}
# Linear Regression with Bootstrap Resampling

set.seed(5)

m1 <- train(lprice ~ points + country,
                  data = wino_tr, 
                  method = "lm", 
                  trControl = trainControl(method = "boot", number = ))

m1
```


```{r}
# RMSE for Test Partition of Data
postResample(
  pred = predict(m1, wino_te),
  obs = wino_te$lprice
)

## RMSE = 0.6138948
```

# Variable selection

We now graph the importance of your 10 features.

```{r}
# Graph!
plot(varImp(m1, scale = TRUE))
```